A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate

### 2020-05-09 Manuscript/SupportingInfo-MLC.Rmd
SI update

## Effect Sizes of Pagel's Lambda ($Z_\lambda$) and Kappa ($Z_K$)

We proposed the use of effect sizes (Z-scores) as a measure of the strength of phylogenetic signal. These were obtained from summary parameters of phylogenetic signal: $\lambda$ and $Kappa$. The precision of effect sizes $Z_\lambda$ and $Z_K$ are summarized in the main text, primarily for simulations where $n=32$. Here we present results from additional simulations.

*Results* Effect sizes for Pagel's $\lambda$ ($Z_\lambda$) scale nonlinearly with input levels of phylogenetic signal, and this pattern is consistent across all tree sizes examined (Figure S5). Further, the precision of $Z_\lambda$ varied widely across input levels of phylogenetic signal, with decidely greater variation (less precision) at stronger input levels of phylogenetic signal (Figure S5). By contrast, effect sizes for *Kappa* ($Z_K$) displayed a linear increase with increasing input levels of phylogenetic signal, which remained consistent across all tree sizes examined (Figure S6). Additionally, the precision of $Z_K$ remained constant across input levels of phylogenetic signal (Figure S6). Together these results imply that $Z_K$ is a more reliable estimate of the strength of phylogenetic signal.

@@

\textbf{Figure S6}. Variation in effect size (Z score) of $K$ across input $\lambda$ values on various tree sizes. \hfill\break

@@

\textbf{Figure S7}. Estimates of phylogenetic signal using $Kappa$ across known lambda inputs on various tree sizes.  Plot shows that while $Kappa$ increases with increasing phylogenetic signal, the precision of this estimate also varies. Thus, conversion to a standardized effect size ($Z_K$) is preferred.

>>>>

## Effect Sizes of Pagel's Lambda ($Z_\lambda$) and Kappa ($Z_{\kappa}$)

We proposed the use of effect sizes (Z-scores) as a measure of the strength of phylogenetic signal. These were obtained from summary parameters of phylogenetic signal: $\lambda$ and $\kappa$. The precision of effect sizes $Z_\lambda$ and $Z_{\kappa}$ are summarized in the main text, primarily for simulations where $n=32$. Here we present results from additional simulations.

*Results* Effect sizes for Pagel's $\lambda$ ($Z_\lambda$) scale nonlinearly with input levels of phylogenetic signal, and this pattern is consistent across all tree sizes examined (Figure S5). Further, the precision of $Z_\lambda$ varied widely across input levels of phylogenetic signal, with decidely greater variation (less precision) at stronger input levels of phylogenetic signal (Figure S5). By contrast, effect sizes for $\kappa$ ($Z_{\kappa}$) displayed a linear increase with increasing input levels of phylogenetic signal, which remained consistent across all tree sizes examined (Figure S6). Additionally, the precision of $Z_{\kappa}$ remained constant across input levels of phylogenetic signal (Figure S6). Together these results imply that $Z_{\kappa}$ is a more reliable estimate of the strength of phylogenetic signal.

@@

\textbf{Figure S6}. Variation in effect size (Z score) of $\kappa$ across input $\lambda$ values on various tree sizes. \hfill\break

@@

\textbf{Figure S7}. Estimates of phylogenetic signal using $\kappa$ across known lambda inputs on various tree sizes.  Plot shows that while $\kappa$ increases with increasing phylogenetic signal, the precision of this estimate also varies. Thus, conversion to a standardized effect size ($Z_{\kappa}$) is preferred.
